\Large \bf Uncovering Bugs in Distributed Storage Systems\\during Testing (not in Production!)

### 2016-01-17 vnext.tex
fixes to the paper to prepare for camera ready

Microsoft Azure Storage is a cloud storage system that provides customers the ability to store seemingly limitless amounts of data. It has grown from tens of petabytes in 2010 to exabytes in 2015, with the total number of objects stored well-exceeding 60 trillion.
%~\cite{greenberg2015keynote}.

@@

One of the many responsibilities of an ExtMgr is to ensure that every extent maintains enough \emph{replicas} in the system. To achieve this, an ExtMgr receives frequent periodic \emph{heartbeat} messages from every EN. The failure of an EN is detected by missing heartbeats. An ExtMgr also receives less frequent, but still periodic, \emph{synchronization reports} from every EN. The sync reports list all the extents (and associated metadata) stored on the EN. Based on these two types of messages, an ExtMgr identifies which ENs have failed, and which extents are affected by the failure and are missing replicas as a result. The ExtMgr then schedules tasks to repair the affected extents and distributes the tasks to the necessary ENs. The ENs then repair the extents from their existing replicas in the system and lazily update the ExtMgr via their next periodic sync reports. All this communication between an ExtMgr and the ENs occurs via network engines installed in each component of vNext (see Figure~\ref{fig:vnext}).

@@

\noindent
Despite of the extensive testing efforts, the vNext developers have been plagued for months by an elusive bug in the ExtMgr logic. All the unit test suites and integration test suites successfully pass every single time. However, the stress test suite fails \emph{from time to time} after very long executions; in these cases, certain replicas of some extents fail and are never repaired.

This bug has proven to be difficult to identify, reproduce and troubleshoot. First, an extent never being repaired is \emph{not} a property that can be easily checked. Second, the bug seems to manifest only in very long executions. Finally, by the time that the bug is detected, very long execution traces have been collected, which makes manual inspection tedious and ineffective.
%To uncover this bug, as well as many other tricky bugs, the vNext developers are in constant search of a generic and systematic approach for testing distributed systems.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{img/modeled_vnext}
\vspace{-5mm}
\caption{Real Extent Manager with a modeled environment (each box represents one \psharp machine).}
\label{fig:azurestoremodel}
\vspace{-2mm}
\end{figure}

@@

  void Init() {
    ExtMgr = new ExtentManager();
    ExtMgr.NetEngine = new ModelNetEngine(); // model network
    ExtMgr.IsModelingTimer = true;	 // disable internal timer
  }

@@

\textbf{Real Extent Manager}.
The real ExtMgr (see Figure~\ref{fig:extentmanager}) contains two data structures: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}. The \texttt{ExtentCenter} maintains the records that map from extents to their hosting ENs. It is updated upon receiving the periodic sync reports from the ENs. Recall that the sync report from a particular EN lists all the extents stored at the EN. Its purpose is to update ExtMgr's possible out-of-date view of the EN with the ground truth. The \texttt{ExtentNodeMap} records the latest heartbeat time from every EN.

@@

\textbf{Intercepting network messages}.
The real ExtMgr uses a network engine to asynchronously send messages to the \texttt{ExtentNode} machines. The \psharp test harness models the original network engine in vNext by overriding its interface. The modeled network engine (see Figure~\ref{fig:enginecode}) intercepts all outbound messages from the ExtMgr, and invokes \texttt{PSharp.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding ENs. This modeled network engine replaces the real network engine in the wrapped ExtMgr (see Figure~\ref{fig:wrap_target}.

@@

// modeled engine for intercepting Extent Manager messages
class ModelNetEngine : NetworkEngine {
  public override void SendMessage(Socket s, Message msg) {
    // intercept and relay Extent Manager messages
    PSharp.Send(TestingDriver, new ExtMgrMsgEvent(), s, msg);
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Modeled network engine in vNext.}
\label{fig:enginecode}
\vspace{-3mm}
\end{figure}

@@

Messages coming from the \texttt{ExtentNode} machines do \emph{not} go through the modeled network engine. They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the wrapped ExtMgr with \texttt{ExtMgr.ProcessMessage} (see Figure~\ref{fig:wrap_target}). The benefit of this approach is that the real ExtMgr can be tested without modifying its code; the ExtMgr is simply unaware of the \psharp test harness and behaves as if it is running in a real distributed environment and communicating with real ENs.

@@

  // extent node failure logic
  [OnEvent(FailureEvent, ProcessFailure)]
  void ProcessFailure() {
    // notifies the monitor that this EN failed
    PSharp.Notify<RepairMonitor>(new ENFailedEvent(), this);
    PSharp.Halt(); // terminate this P# machine
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Code snippet of the modeled EN in vNext.}
\label{fig:modeled_en}
\vspace{-2mm}
\end{figure}

@@

System correctness should \emph{not} hinge on the frequency of any individual timer. Hence, it makes sense to delegate all nondeterminism due to timing-related events to \psharp. To achieve this, all the timers inside ExtMgr are disabled (see Figure~\ref{fig:wrap_target}), and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp.
%, an approach also used in previous work~\cite{desai2015building}.
Similarly, \texttt{ExtentNode} machines do {\em not} have internal timers either. Their periodic heartbeats and sync reports are also driven by timers modeled in \psharp.

@@

The original \psharp trace did not include sufficient details to allow the developers to identify the root cause of the problem. Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debugging outputs in each iteration. After several iterations, the developers were able to pinpoint the exact culprit and immediately propose a solution for fixing the bug. Once the proposed solution was implemented, the developers ran the test harness again. No bugs were found for 100,000 executions, a process that only took on the order of tens of minutes.

>>>>

Microsoft Azure Storage is a cloud storage system that provides customers the ability to store seemingly limitless amounts of data. It has grown from tens of petabytes in 2010 to exabytes in 2015, with the total number of objects stored well-exceeding 60 trillion~\cite{greenberg2015keynote}.

@@

One of the responsibilities of an ExtMgr is to ensure that every managed extent maintains enough \emph{replicas} in the system. To achieve this, an ExtMgr receives frequent periodic \emph{heartbeat} messages from every EN that it manages. EN failure is detected by missing heartbeats. An ExtMgr also receives less frequent, but still periodic, \emph{synchronization reports} from every EN that it manages. The sync reports list all the extents (and associated metadata) stored on the EN. Based on these two types of messages, an ExtMgr identifies which ENs have failed, and which extents are affected by the failure and are missing replicas as a result. The ExtMgr then schedules tasks to repair the affected extents and distributes the tasks to the necessary ENs. The ENs then repair the extents from the existing replicas and lazily update the ExtMgr via their next periodic sync reports. All this communication between an ExtMgr and the ENs occurs via network engines installed in each component of vNext (see Figure~\ref{fig:vnext}).

@@

\noindent
Despite of the extensive testing efforts, the vNext developers were plagued for months by an elusive bug in the ExtMgr logic. All the unit test suites and integration test suites successfully passed on each test run. However, the stress test suite failed \emph{from time to time} after very long executions; in these cases, certain replicas of some extents failed without subsequently being repaired.

This bug proved difficult to identify, reproduce and troubleshoot. First, an extent never being repaired is \emph{not} a property that can be easily checked. Second, the bug appeared to manifest only in very long executions. Finally, by the time that the bug did manifest, very long execution traces had been collected, which made manual inspection tedious and ineffective.
%To uncover this bug, as well as many other tricky bugs, the vNext developers are in constant search of a generic and systematic approach for testing distributed systems.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{img/modeled_vnext}
\vspace{-5mm}
\caption{Real Extent Manager with its \psharp test harness (each box represents one \psharp machine).}
\label{fig:azurestoremodel}
\vspace{-2mm}
\end{figure}

@@

  void Init() {
    ExtMgr = new ExtentManager();
    ExtMgr.NetEngine = new ModelNetEngine(); // model network
    ExtMgr.DisableTimer();	 // disable internal timer
  }

@@

\textbf{Real Extent Manager}.
The real ExtMgr (see Figure~\ref{fig:extentmanager}) contains two data structures: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}. The \texttt{ExtentCenter} maps extents to their hosting ENs. It is updated upon receiving a periodic sync reports from an EN. Recall that a sync report from a particular EN lists all the extents stored at the EN. Its purpose is to update the ExtMgr's possible out-of-date view of the EN with the ground truth. The \texttt{ExtentNodeMap} maps ENs to their latest heartbeat times.

@@

\textbf{Intercepting network messages}.
The real ExtMgr uses a network engine to asynchronously send messages to ENs. The \psharp test harness models the original network engine in vNext by overriding the original implementation. The modeled network engine (see Figure~\ref{fig:enginecode}) intercepts all outbound messages from the ExtMgr, and invokes \texttt{PSharp.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding ENs. This modeled network engine replaces the real network engine in the wrapped ExtMgr (see Figure~\ref{fig:wrap_target}).

@@

// modeled engine for intercepting Extent Manager messages
class ModelNetEngine : NetworkEngine {
  public override void SendMessage(Socket s, Message msg) {
    // intercept and relay Extent Manager messages
    PSharp.Send(TestingDriver, new ExtMgrMsgEvent(), s, msg);
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Modeled vNext network engine.}
\label{fig:enginecode}
\vspace{-3mm}
\end{figure}

@@

Messages coming from \texttt{ExtentNode} machines do \emph{not} go through the modeled network engine; they are instead delivered to the \texttt{ExtentManager} machine and trigger an action that invokes the messages on the wrapped ExtMgr with \texttt{ExtMgr.ProcessMessage} (see Figure~\ref{fig:wrap_target}). The benefit of this approach is that the real ExtMgr can be tested without modifying the original communication-related code; the ExtMgr is simply unaware of the \psharp test harness and behaves as if it is running in a real distributed environment and communicating with real ENs.

@@

  // extent node failure logic
  [OnEvent(FailureEvent, ProcessFailure)]
  void ProcessFailure() {
    // notifies the monitor that this EN failed
    PSharp.Notify<RepairMonitor>(new ENFailedEvent(), this);
    PSharp.Halt(); // terminate this P# machine
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Code snippet of the modeled EN.}
\label{fig:modeled_en}
\vspace{-2mm}
\end{figure}

@@

System correctness should \emph{not} hinge on the frequency of any individual timer. Hence, it makes sense to delegate all nondeterminism due to timing-related events to \psharp. To achieve this, all the timers inside ExtMgr are disabled (see Figure~\ref{fig:wrap_target}), and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp, an approach also used in previous work~\cite{desai2015building}.\footnote{We had to make a minor change to the real ExtMgr code to facilitate modeling: we added the \texttt{DisableTimer} method, which disables the real ExtMgr timer so that it can be replaced with the \psharp timer.}
Similarly, \texttt{ExtentNode} machines do {\em not} have internal timers either. Their periodic heartbeats and sync reports are also driven by timers modeled in \psharp.

@@

The original \psharp trace did not include sufficient details to allow the developers to identify the root cause of the problem. Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debugging outputs in each iteration. After several iterations, the developers were able to pinpoint the exact culprit and immediately propose a solution for fixing the bug. Once the proposed solution was implemented, the developers ran the test harness again. No bugs were found during 100,000 executions, a process that only took on the order of tens of minutes.
